---

author:  Regression
date: 06 May, 2021
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    
---

<br>

In this task, The famous Boston dataset from the MASS package records the historical median house value (medv column, in 1000s USD) for 506 suburbs around Boston in the 1970s.

<br>

# Task

<br>


####  LOADING THE DATASET 
```{r}
# call install.packages("MASS") first (only once)
library("MASS")
head(Boston, 3)
```
<br>


<br><br>

#### 1. Fit a simple linear model of medv as a function of lstat.
First loading packages,
```{r echo=T, results='hide',error=FALSE,warning=FALSE,message=FALSE}
f_linear <- lm(medv~lstat, data=Boston)
```


<br><br>

#### 2. Fit a quadratic polynomial model (ğ‘ğ‘‹^2 + ğ‘ğ‘‹ + ğ‘) for the same pair of vari- ables.
```{r}
f_quad <- lm(medv ~ poly(lstat, 2, raw=TRUE), data=Boston)

```

<br><br>

#### 3. Fit a cubic polynomial model (ğ‘ğ‘‹^3 + ğ‘ğ‘‹^2 + ğ‘ğ‘‹ + ğ‘‘) for the same pair of variables.

```{r}
f_cubic <- lm(medv ~ poly(lstat, 3, raw=TRUE), data=Boston)
```

<br><br>

#### 4. Draw the scatter plot of the two variables and add the fitted regression curves (all three on a single plot, use different colours).

First,getting test values of `lstat` in `test` 
```{r echo=T,error=FALSE,warning=FALSE,message=FALSE}
test <- seq(min(Boston$lstat, na.rm=TRUE),
            max(Boston$lstat, na.rm=TRUE),
            length.out=201)
```
now, getting desired output from our fitted models 
```{r}
yy_linear <- f_linear$coefficients[1] + f_linear$coefficients[2]*(test)
yy_quad   <- f_quad$coefficients[1] + f_quad$coefficients[2]*(test) +      
             f_quad$coefficients[3]*(test^2)

yy_cubic  <- f_cubic$coefficients[1] + f_cubic$coefficients[2]*(test) +    
             f_cubic$coefficients[3]*(test^2) + 
             f_cubic$coefficients[4]*(test^3)
```
Finally, plotting 
```{r}
plot(medv~lstat,data=Boston, las=1,col=c(5,7),  lwd=1)
lines(test, yy_linear, col=2, lwd=3)
lines(test, yy_quad, col=3, lwd=3)
lines(test, yy_cubic, col=4, lwd=3)
legend(25, 45,legend=c("linear", "quad","cubic"), col=2:4, lty="solid")



```

<br><br>

#### 5. Construct a multiple regression model for medv as a function of lstat, rm, and tax.
```{r}
f_mul <- lm(medv~lstat + rm + tax, data=Boston)
```

<Br><br>

#### 6. Use for ward selection (with respect to the AIC criterion) to come up with a multiple regression model for medv as a function of other variables.
```{r echo=T,error=FALSE,warning=FALSE,message=FALSE }
model_full <- formula(model.frame(medv~., data=Boston))
model_empty <- medv ~ 1

step(f_mul, scope=model_full, direction="forward")
```
hence, we get our desired variables for best fitting model. With best fitting variables i.e `indus` and `age`. Also "none" gives lowest value of AIC too. 
<br><br>

#### 7. Use backward elimination (with respect to the AIC criterion) to construct a multiple regression model for medv as a function of other variables.

```{r echo=T,error=FALSE,warning=FALSE,message=FALSE}
step(lm(model_full, data=Boston), scope=model_empty, direction="backward")
```

 
 <br><br>

#### 8. Construct a multiple regression model for medv as a function of lstat, rm, and tax transformed in various ways. Apply logarithms, squares, exponen- tial (amongst others) functions on the variables and try to come up with the best model by trial and error (you can also use forward selection and/or back- ward elimination for this).

Constructing medv as a function of lstat, rm, and tax transformed in various ways.
```{r}
f_mul <-   lm(medv~lstat + rm + tax, data=Boston)

f_poly <-  lm(medv~ poly(lstat, 4, raw=TRUE) + poly(rm, 4, raw=TRUE) +
           poly(tax, 4, raw=TRUE), data=Boston)

f_log <-   lm(medv ~ log(lstat) + log(rm) + log(tax), data=Boston)
f_squ <-   lm(medv ~ lstat^2 + rm^2 + tax^2, data=Boston)
f_expo <-  lm(medv ~ exp(lstat) + rm + tax, data=Boston)

model_empty <- formula(medv ~ 1) 

x_full <- formula( medv ~  lstat+ rm + tax + poly(lstat, 3, raw=TRUE) +
          poly(rm, 3, raw=TRUE) + poly(tax, 4, raw=TRUE) +
          log(lstat) + log(rm) + log(tax) + lstat^2 + rm^2 + 
          tax^2 + exp(lstat) + exp(rm) 
           )
```

```{r}
step(lm(medv~ 1, data=Boston,  na.action=na.exclude), scope=x_full,
     direction="forward")

```
finally, applying `step`,  <br>
"exp(lstat) + log(tax)" has lowest AIC, so these are best variables for this model . 


<br>
<br>  



#### 9. Compute AIC, RMSE, MAE, and adjusted ğ‘…2 of each model. Which model is the best with regards to each metric? Draw conclusions in your own words.

```{r}
# r square 
r_squ <- data.frame(RSquare=" ", Linear=summary(f_linear)$adj.r.squared,  
         Quad= summary(f_quad)$adj.r.squared,   
         Cubic=summary(f_cubic)$adj.r.squared 
            )

#MAE
MAE  <- data.frame(MAE=" ", Linear=mean(abs(f_linear$residuals)),
        Quad=mean(abs(f_quad$residuals)),
        Cubic= mean(abs(f_cubic$residuals))
             )

#RMSE
RMSE <- data.frame(RMSE=" ",Linear=sqrt(mean(f_linear$residuals^2)),
        Quad=sqrt(mean(f_quad$residuals^2)),
        Cubic=sqrt(mean(f_cubic$residuals^2))
             )

#AIC
AIC <-  data.frame(AIC=" ", linear=AIC(f_linear),
        Quad= AIC(f_quad),
        Cubic=AIC(f_cubic)
            )
```
```{r}
knitr::kable(r_squ) 
knitr::kable(MAE)
knitr::kable(RMSE)
knitr::kable(AIC)
```
As, taking AIC in account, we can conclude that Cubic model is more precise and have more accurecy. 



<br> <br> 

#### 10. For each model, draw the plot of the residuals (ğ‘¦Ì‚ âˆ’ ğ‘¦ ) as a function of the predicted outputs (ğ‘¦Ì‚ ). Describe these plots in your own words.
```{r echo=T, results='hide',error=FALSE,warning=FALSE,message=FALSE}
library(tidyverse)
library(ggpubr)

f_linear <- lm(medv~lstat, data=Boston)

par(mfrow=c(2, 2))
p1  <- ggplot(data=Boston,
       mapping=aes(x=predict(f_linear), y= f_linear$residuals  ))  +
       geom_point(col="red") + labs(title="Residuals vs predicted (Linear model)",   
       x="predicted medv outputs",y="residuals")

p2 <-ggplot(data=Boston,
      mapping=aes(x=predict(f_quad), y= f_quad$residuals  ))  +    
      geom_point(col="red") + labs(title="Residuals vs predicted(Quad model)",
      x="predicted medv outputs",y="residuals")                            

p3 <-ggplot(data=Boston,
      mapping=aes(x=predict(f_cubic), y= f_cubic$residuals  ))  + 
      geom_point(col="red") + labs(title="Residuals vs predicted(Cubic model)", 
      x="predicted medv outputs",y="residuals")                            

ggarrange(p1, p2, p3, ncol=2, nrow=2) 

```
The interpretation of a "residuals vs. predictor plot" is identical to that for a "residuals vs. fits plot." That is, a well-behaved plot will bounce randomly and form a roughly horizontal band around the residual = 0 line. And, no data points will stand out from the basic random pattern of the other residuals. <br> 
if there is some non-random pattern to the plot, it indicates that it would be worthwhile adding the predictor to the model. <br>
Linear model or Cubic  :The points on the plot show no pattern or trend, suggesting that there is no relationship between the residuals and medv.  <br>
Quad model : well, it seems a negative slope pattern or trend , so "lstat" variable doing good in it. <br>


<br> <br>

#### 11. Predict the medv values for lstat of 0, 25, 50, and 75 using all the models. Compare and discuss the results. Which of the predictions seem trustwor- thy?

```{r echo=T,error=FALSE,warning=FALSE,message=FALSE}

test <- data.frame( lstat=c(0,25,50,75))
values <- data.frame( lstat=c(0,25,50,75), 
                      linear_pred_medv =predict(f_linear,test),                                             Quad_pred_medv = predict(f_quad, test), 
                      cubic_pred_medv = predict(f_cubic, test)  
                      )
knitr::kable(values)
```

<br>



<br><br><br><br>



